MIMOZA

### 2017-09-21 MIMOZA/mimoza.Rmd
added hydroex

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

@@

```{r}
if (!file.exists('mimoza.imilast.rda')) {
  imilast.storms <- read.imilast('~/data/IMILAST/ERAinterim_0.75_SH_M07_19790101_20091231.txt',lon=c(30,90),lat=c(-50,-5))
  cyclones <- events2trajectory(imilast.storms)
  save(cyclones,file='mimoza.imilast.rda')
} load('mimoza.imilast.rda')
```

@@

```{r}
## get the GHCND data
if (!file.exists('mimoza.ghcnd.rda')) {
  ## Daily maximum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmax')
  tmax <- station(ss,verbose=TRUE)
  ## Daily minimum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmin')
  tmin <- station(ss,verbose=TRUE)
  ## Daily precipitation
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='precip')
  precip <- station(ss,verbose=TRUE)
  ## Set bad (missing) data to NA
  z <- coredata(tmax); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmax)
  z <- coredata(tmin); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmin)
  z <- coredata(precip); z[z < 0] <- NA; z[z > 700] <- NA; z -> coredata(precip)
  ## Save in R-binary for convenience.
  save(tmax,tmin,precip,file='mimoza.ghcnd.rda')
} else load ('mimoza.ghcnd.rda')
```

>>>>

## Introduction.

This is an R Markdown document that explores some climate data with relevance to Mozambique. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

@@

```{r}
if (!file.exists('mimoza.imilast.rda')) {
  imilast.storms <- subset(read.imilast('~/data/IMILAST/ERAinterim_0.75_SH_M07_19790101_20091231.txt'),
                           is=list(lon=c(20,90),lat=c(-40,-5)))
# imilast.storms <- subset(read.imilast('~/data/IMILAST/ERAinterim_1.5_SH_M03_19890101_20090331_ST.txt'),
#                           is=list(lon=c(30,90),lat=c(-40,-5)))
  cyclones <- events2trajectory(imilast.storms)
  save(cyclones,file='mimoza.imilast.rda')
} else load('mimoza.imilast.rda')
ii <- is.element(colnames(cyclones),'amp_z850')
mi <- apply(cyclones[,ii],1,min)
storms <- cyclones[mi < 10,]
storms <- attrcp(cyclones,storms)
class(storms) <- class(cyclones)
col <- rgb(0,0,1,1-(mi-min(mi))/diff(range(mi)))
map(storms,verbose=TRUE,lonR=30,latR=0,col=col,lwd=3)
```

```{r}
pca.storms <- PCA(storms,verbose=TRUE)
plot(pca.storms[,1],type='l')
plot(attr(pca.storms,'pattern')[,1],type='l')
plot(attr(pca.storms,'eigenvalues'),type='b')
```

@@

```{r}
## get the GHCND data
if (!file.exists('mimoza.ghcnd.rda')) {
  ## Daily maximum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmax')
  tmax <- station(ss,verbose=TRUE)
  ## Daily minimum temperature
  ss <- select.station(lon=c(27,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='tmin')
  tmin <- station(ss,verbose=TRUE)
  ## Daily precipitation
  ss <- select.station(lon=c(31,37),lat=c(-40,-10),src='ghcnd',nmin=30,param='precip')
  precip <- station(ss,verbose=TRUE)
  ## Set bad (missing) data to NA
  z <- coredata(tmax); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmax)
  z <- coredata(tmin); z[z < -20] <- NA; z[z > 70] <- NA; z -> coredata(tmin)
  z <- coredata(precip); z[z < 0] <- NA; z[z > 700] <- NA; z -> coredata(precip)
  ## Save in R-binary for convenience.
  save(tmax,tmin,precip,file='mimoza.ghcnd.rda')
} else load ('mimoza.ghcnd.rda')
```

Maximum temperature - mean seasonal cycle

```{r}
plot(aggregate(tmax,month),new=FALSE)
plot(aggregate(tmax,month,FUN='sd'),new=FALSE)
```

Minimum temperature - mean seasonal cycle

```{r}
plot(aggregate(tmin,month),new=FALSE)
plot(aggregate(tmin,month,FUN='sd'),new=FALSE)
```

Precipitation - mean seasonal cycle

```{r}
plot(aggregate(precip,month,FUN='sum'),new=FALSE)
plot(aggregate(precip,month,FUN='wetfreq'),new=FALSE)
plot(aggregate(precip,month,FUN='wetmean'),new=FALSE)
```

Precipitation - long-term variations

```{r}
plot(annual(precip,FUN='sum'),new=FALSE); grid()
plot(annual(precip,FUN='wetfreq'),new=FALSE); grid()
plot(annual(precip,FUN='wetmean'),new=FALSE); grid()
```

Spatial patterns of variability in annual precipitation totals

```{r, fig.height=8}
Y <- subset(annual(precip,FUN='sum'),it=c(1950,1990))
nv <- apply(coredata(Y),2,'nv')
Y <- subset(Y,is=nv > 36)
Y <- pcafill(Y)
tp.pca<- PCA(Y)
plot(tp.pca,new=FALSE)
```

Spatial patterns of variability in annual wet-day frequency

```{r, fig.height=8}
Y <- subset(annual(precip,FUN='wetfreq'),it=c(1950,1990))
nv <- apply(coredata(Y),2,'nv')
Y <- subset(Y,is=nv > 36)
Y <- pcafill(Y)
fw.pca<- PCA(Y)
plot(fw.pca,new=FALSE)
```

Spatial patterns of variability in annual wet-day mean precipitation

```{r, fig.height=8}
Y <- subset(annual(precip,FUN='wetmean'),it=c(1950,1990))
nv <- apply(coredata(Y),2,'nv')
Y <- subset(Y,is=nv > 36)
Y <- pcafill(Y)
mu.pca<- PCA(Y)
plot(mu.pca,new=FALSE)
```

## Copernicus land-cover information - 100m resolution over Africa

<http://land.copernicus.eu/global/products/lc>
